# Data Exploration

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(haven)
library(table1)
library(lubridate)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(ggfortify)
library(corrplot)


#library(plotly)

#library(tidycmprsk)
#library(condSURV)
```
The JHS Analysis Dataset Warehouse dictionary was reviewed and variables were selected for further analysis/use in a survival analysis based on the CDC's top risk factors which include: high blood pressure, high cholesterol, smoking, obesity and diabetes. Additionally, age is included, as risk increases with age.

## Summary statistics

Table 2 provides descriptive statistics about selected variables based on the classification of stroke versus the adjudication of stroke and the stroke type. 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
#load visit 1 data
datav1 <- read_dta("data/analysis1.dta")
visit1 <- datav1 %>% dplyr :: select(subjid, visitdate, age, htn, diab3cat, idealhealthsmk, bmi3cat, totchol3cat)

#visit1 <- visit1 %>% rename(visitdate_v1 = visitdate)

#load adjudicated data

#data_allevtstroke <- read_sas("data/allevtstroke.sas7bdat", NULL)
#allevtstroke <- data_allevtstroke %>% dplyr :: select(subjid, Stroke, OHDStroke, eventdate)

data_incevtstroke <- read_sas("data/incevtstroke.sas7bdat", NULL)
incevtstroke <- data_incevtstroke %>% dplyr :: select(subjid, stroke, date, contactType)

#allevtstroke <- allevtstroke %>% rename(stroke_allevt = Stroke, ohdstroke = OHDStroke)

incevtstroke <- incevtstroke %>% rename(contacttype = contactType)

#data_adj <- full_join(allevtstroke, incevtstroke, by = "subjid")

# Create final data set

data <- full_join(incevtstroke, visit1, by = "subjid")
data <- na.omit(data)

#Transform data
data$stroke <- as.factor(data$stroke)
data$htn <- as.factor(data$htn)
data$diab3cat <- as.factor(data$diab3cat)
data$idealhealthsmk <- as.factor(data$idealhealthsmk)
data$totchol3cat <- as.factor(data$totchol3cat)
data$bmi3cat <- as.factor(data$bmi3cat)

##Label data
label(data$stroke) <- "Adjudicated Stroke Occurrance"
levels(data$stroke) <- c("No Stroke", "Adjudicated Stroke")

label(data$htn) <- "Hypertensive Status"
levels(data$htn) <- c("Normal/Pre-Hypertensive", "Hypertensive")

label(data$diab3cat) <- "Diabetes Categorization"
levels(data$diab3cat) <- c("Non-Diabetic", "Pre-Diabetic", "Diabetic")

label(data$idealhealthsmk) <- "Smoking Status"
levels(data$idealhealthsmk) <- c("Current Smoker","Never Smoked/Quit + 12 months")

label(data$bmi3cat) <- "AHA BMI Categorization"
levels(data$bmi3cat) <- c("Obese","Overweight","Normal Weight")

label(data$totchol3cat) <- "AHA Total Cholesterol Categorization"
levels(data$totchol3cat) <- c("Poor Health","Intermediate Health","Ideal Health")

#Create Table
#table1(~ stroke + contacttype + age + htn + diab3cat + idealhealthsmk + bmi3cat + totchol3cat, data = data)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}

#Create time variable (time from visit 1 until event/censoring date)
data <- data %>% mutate(time = difftime(date, visitdate))

#Create the survival object (created from time variable and if the event has taken place.)
obj <- Surv(data$time, data$stroke)
```


### **Summary Statistics Based on Contact Type** {.unnumbered}
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Create Table based on contact type
table1(~ age + htn + diab3cat + idealhealthsmk + bmi3cat + totchol3cat|contacttype, data=data)

```

### **Visuals** {.unnumbered}
```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Create plot of smoking vs contact type.

plot1 <- data %>% ggplot(aes(colour=as.factor(idealhealthsmk), x=contacttype)) + geom_bar (method=lm) + labs(colour= "Smoking Status", y = "Number of Participants", x = "Contact type", title = "Contact type and Smoking Status")

plot1

```

Note: Looking at the above plot it can be seen that the majority of current smokers were censored from the study.



The below plots illustrate how the individual variables trend (and change) throughout the study.

```{r, echo = FALSE, warning = FALSE, message = FALSE}

#load visit 1 data
datav1 <- read_dta("data/analysis1.dta")
visit1 <- datav1 %>% dplyr :: select(subjid, visitdate, age, htn, diab3cat, idealhealthsmk, bmi3cat, totchol3cat)

#visit1 <- visit1 %>% rename(visitdate_v1 = visitdate)

#load adjudicated data

#data_allevtstroke <- read_sas("data/allevtstroke.sas7bdat", NULL)
#allevtstroke <- data_allevtstroke %>% dplyr :: select(subjid, Stroke, OHDStroke, eventdate)

data_incevtstroke <- read_sas("data/incevtstroke.sas7bdat", NULL)
incevtstroke <- data_incevtstroke %>% dplyr :: select(subjid, stroke, date, contactType)

#allevtstroke <- allevtstroke %>% rename(stroke_allevt = Stroke, ohdstroke = OHDStroke)

incevtstroke <- incevtstroke %>% rename(contacttype = contactType)

#data_adj <- full_join(allevtstroke, incevtstroke, by = "subjid")

# Create final data set

data <- full_join(incevtstroke, visit1, by = "subjid")
data <- na.omit(data)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
#Create time variable (time from visit 1 until event/censoring date)
data <- data %>% mutate(time = difftime(date, visitdate))

#Create the survival object (created from time variable and if the event has taken place.)
obj <- Surv(data$time, data$stroke)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}

##Label data
label(data$age) <- "Age"
label(data$stroke) <- "Adjudicated Stroke Occurrance"
label(data$htn) <- "Hypertensive Status"
label(data$diab3cat) <- "Diabetes Categorization"
label(data$idealhealthsmk) <- "Smoking Status"
label(data$bmi3cat) <- "AHA BMI Categorization"
label(data$totchol3cat) <- "AHA Total Cholesterol Categorization"
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
#fit <- survfit(Surv(data$time, data$stroke) ~ data$age)
#autoplot(survfit(Surv(data$time, data$stroke) ~ 1),surv.colour = 'orange', censor.colour = 'red')

# https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_surv.html
# https://www.rdocumentation.org/packages/survival/versions/3.4-0/topics/aareg
# https://rviews.rstudio.com/2017/09/25/survival-analysis-with-r/


plot2 <- autoplot(aareg(obj ~ data$age + data$htn + data$diab3cat + data$idealhealthsmk + data$bmi3cat + data$totchol3cat))

#plot2 <- plot2 %>% label()

plot2
```

## Data Dictionary

Two data sets were utilized and joined to create a final data set for analysis. Following are the variables used.

The time variable that will be utilized in the survival analysis is calculated using V1DATE and DATE from the adjudicated data set (incevtstroke). This calculated variable represents the event time which is the time from a participant joining the study until the participant either has a stroke or is censored from the study.


### **DATA SET:** Visit 1 Data (analysis1) {.unnumbered}

These data include observations from the participant's first visit in the study.

| **Variable Name** | **Definitions and Attributes**                                                                                                                                                                                                                                                                                                      |
|-------------------|-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| SUBJID            | - Subject identification number for participant                                                                                                                                                                                                                                                                                     |
| VISITDATE         | - Date of participant exam visit                                                                                                                                                                                                                                                                                                    |
| AGE               | - Age (in years) of participant at time of visit 1 <br> - Age, based on birth date and clinic visit date, rounded to 0.1                                                                                                                                                                                                                 |
| HTN               | - HTN status is defined as blood pressure $\geq$ 140/90 mmHg (per JNC 7) or use of blood pressure lowering medication <br> - 0 = if BPjnc7 $\leq$ 1 (Normal/Pre-Hypertensive) AND participant NOT on medication <br> - 1 = if BPjnc7 $\geq$ 2 (Stage I/II Hypertension) AND participant uses medication to control hypertension     |
| DIAB3CAT          | **Diabetes categorization** <br> - 0 = Non-Diabetic <br> - 1 = Pre-Diabetic <br> - 2 = Diabetic**                                                                                                                                                                                                                                   |
| IDEALHEALTHSMK    | **Indicator for Ideal Health via Smoking Status** <br> - 0 = Current smoker or Quit \< 12 months ago <br> - 1 = Never smoked or Quit $\geq$ 12 months ago                                                                                                                                                                           |
| BMI3CAT           | - 0 = Poor Health defined as BMI $\geq$ 30 (Obese) <br> - 1 = Intermediate Health defined as 25 $\leq$ BMI \< 30 (Overweight) <br> - 2 = Ideal Health defined as BMI \< 25 (Normal)                                                                                                                                                 |
| TOTCHOL3CAT       | **AHA Total Cholesterol Categorization** <br> - 0 = Poor Health defined as Total Cholesterol $\geq$ 240 <br> - 1 = Intermediate Health defined as 200 $\leq$ Total Cholesterol \< 240 OR if Total Cholesterol \< 200 as controlled by medication <br> - 2 = Ideal Health defined as Total Cholesterol \< 200 with NO medication use |


### **DATA SET:** Adjudicated Data (incevtstroke) {.unnumbered}


| **Variable Name** | **Definitions and Attributes**                                                                                                         |
|-------------------|----------------------------------------------------------------------------------------------------------------------------------------|
| SUBJID            | - Subject identification number for participant                                                                                        |
| STROKE            | - 0 = No stroke <br> - 1 = Adjudication of stroke                                                                                                   |
| V1DATE            | Date of Exam 1 Clinic Visit                                                                                                            |
| DATE              | Event Date (stroke) or Censoring Date                                                                                                  |
| CONTACTTYPE       | ** Last Contact Type ** <br> - Censored <br> - Hemorrhagic Stroke <br> - Ischemic Stroke <br> - Previous Stroke History <br> - Refused |
