# Methods

Data reported by the JHS study was used in calculating basic survival analysis on the adjudication of strokes in the cohort. Two data sets were utilized and joined on the variable subjectid. After removing N/As, 2,288 observations remained in the data set for further analysis. The analysis was completed using R statistical software package, specifically the {survival} package. A time variable was created from the date at visit 1 until either the date of event (stroke) or the censoring date to use in the analysis. 


## Creating a survival object
Using the calculated time and the variable stroke, a survival object was created using the surv function. The object creates a vector that results in one entry created for each observation that represent the time to stroke. If the subject was censored, the entry is followed by a +. A sample of the object vector is shown:


```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(haven)
library(table1)
library(lubridate)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(ggfortify)
library(corrplot)
library(fastDummies)

#library(tidycmprsk)
#library(condSURV)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}

#load visit 1 data
datav1 <- read_dta("data/analysis1.dta")
visit1 <- datav1 %>% dplyr :: select(subjid, visitdate, age, htn, diab3cat, idealhealthsmk, bmi3cat, totchol3cat)

#visit1 <- visit1 %>% rename(visitdate_v1 = visitdate)

#load adjudicated data

#data_allevtstroke <- read_sas("data/allevtstroke.sas7bdat", NULL)
#allevtstroke <- data_allevtstroke %>% dplyr :: select(subjid, Stroke, OHDStroke, eventdate)

data_incevtstroke <- read_sas("data/incevtstroke.sas7bdat", NULL)
incevtstroke <- data_incevtstroke %>% dplyr :: select(subjid, stroke, date, contactType)

#allevtstroke <- allevtstroke %>% rename(stroke_allevt = Stroke, ohdstroke = OHDStroke)

incevtstroke <- incevtstroke %>% rename(contacttype = contactType)

#data_adj <- full_join(allevtstroke, incevtstroke, by = "subjid")

# Create final data set

data <- full_join(incevtstroke, visit1, by = "subjid")
data <- na.omit(data)

```



```{r, echo = FALSE, warning = FALSE, message = FALSE}

#Create time variable (time from visit 1 until event/censoring date)
data <- data %>% mutate(time = difftime(date, visitdate))

#Create the survival object (created from time variable and if the event has taken place.)
obj <- Surv(data$time, data$stroke)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
 #Sample from object vector
 tail(head(obj, 115), 5)
```


The object created was then used to construct a survival curve using the survfit function. As an example, the object was plotted against hypertension and diabetes categorization. 

**Hypertension**

*
As an example, the below graph illustrates the probability of **NOT** having a stroke (the event):
  + Non-hypertensive (pink): approximately 1% decrease over time
  + Hypertensive (blue): approximately 3% decrease over time
```{r, echo = FALSE, warning = FALSE, message = FALSE}
# https://cran.r-project.org/web/packages/ggfortify/vignettes/basics.html

autoplot(survfit(obj ~ htn, data = data))
```

**Diabetes**

*
As an example, the below graph illustrates the probability of **NOT** having a stroke (the event):
  + Non-Diabetic (pink): < 1% decrease over time 
  + Pre-Diabetic (green): approximately 2.5% decrease over time
  + Diabetic (blue): approximately 4.5% decrease over time

```{r, echo = FALSE, warning = FALSE, message = FALSE}
autoplot(survfit(obj ~ diab3cat, data = data))

```



## Kaplan-Meier (K-M) Survival Estimate {.unnumbered}

Broken down, the K-M survival probability of not having stroke at time, $t_j$, $S(t_j)$ can be calculated using: 

+ Formula variable definitions

  + $S(t_{j-1})$ => the probability of not having a stroke at time *$t_{j-1}$
  + $n_j$ => the number of participants stroke free just before $t_j$, $n_j$
  + $d_j$ => the number of events (strokes) at $t_j$
  
  $S(t_j)$ = $S(t_{j-1})$(1 - $\frac{d_j}{n_j}$)


A Kaplan-Meier plot was created to visualize the survival curve and confidence interval, which is further discussed in Data Analysis and Results.


## Cox Proportional Hazard (PH) {.unnumbered}

The Cox PH can is the hazard function, $h(t)$, which is determined by the baseline hazard ($h_0(t)$), a set of variables ($x_1, x_2, … x_p$), and the effect of each variable is determined by their coefficient ($b_1, b_2 … b_p$) (11), mathematically shown as:

$$
h(t) = h_0(t) * e^{b_1x_1 + b_2x_2 + … + b_px_p}
$$

This method does not make the assumption that survival times follow any type of statistical distribution. In the development of the model, it is important to consider whether colinearity, the correlation between two of the independent variables, exists (14). 

A Cox PH analysis was used to model multiple variables from the data set. The coxph() function from the {survival} package was used. The object previously created for the KM plot is found on the left side with standard regression formulas on the right. All variables considered in the model are identified by the CDC as major risk factors for stroke. A table of results was created using the tbl_regression() function out of {gtsummary}, with exponentiate set to TRUE to return the hazard ratio versus the log hazard ratio, which is further discussed in Data Analysis and Results.



## Assumption Verification {.unnumbered}

**Correlation Plot Of Stroke Risk Data**
```{r, echo = FALSE, warning = FALSE, message = FALSE}
##Colinearity/Multi-collinearity?? 

data_sub <- data %>% dplyr :: select(stroke, age, htn, diab3cat, idealhealthsmk, bmi3cat, totchol3cat)

A <- cor(data_sub)
colnames(A) = c('chol', 'BMI','Stroke','age','HTN','Dia','Smoke')

corrplot.mixed(A, order = 'AOE')
```

**Lack of Fit**
```{r, echo = FALSE, warning = FALSE, message = FALSE}
##LOF


```

**Proportional Hazards**
Assumption: Hazards are proportional at each point in time throughout the followup
Using the cox.zph() function from the {survival} package, the assumption can be checked.
STILL WORKING ON CODE

```{r, echo = FALSE, warning = FALSE, message = FALSE}


assump_fit <- coxph(obj ~ age + as.factor(htn) + as.factor(diab3cat) + as.factor(idealhealthsmk) + as.factor(bmi3cat)+ as.factor(totchol3cat), data=data)
#assump_fit


# Checking PH assumptions
# Significat p-value indicates assumption is not met
mv_fit <- coxph(obj ~ stroke + age + htn+ diab3cat + idealhealthsmk + bmi3cat + totchol3cat, data = data)
cz <- cox.zph(mv_fit)
print(cz)
plot(cz)

```









